Automated Residual Plot Assessment with the R Package autovi and the Shiny App autovi.web

### 2025-05-12 autovi_paper.qmd
update address and email

---
title: "Automated Residual Plot Assessment with the R Package autovi and the Shiny App autovi.web"
format:
  quarto-anzjs-pdf:
    keep-tex: true  
    shorttitle: "Automated Residual Plot Assessment"
    double-space: true
    line-numbers: true
    # You can specify extra preamble in the file below.
    include-in-header:
      - header.tex
date: today
date-format: "YYYY" # used to define volumeyear
author:
  - name: Weihao Li
    email: weihao.li@monash.edu
    corresponding: true
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0003-4959-106X
  - name: Dianne Cook
    email: dicook@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0002-3813-7155
  - name: Emi Tanaka
    email: emi.tanaka@anu.edu.au
    affiliations: 
      - name: The Australian National University
        department: Biological Data Science Institute
        address: 46 Sullivan's Creek Road, ACT 2600, Australia
    orcid: 0000-0002-1455-259X
  - name: Susan VanderPlas
    email: emi.tanaka@anu.edu.au
    affiliations:
      - name: University of Nebraska
        department: Department of Statistics
        address: Hardin Hall, 3310 Holdrege St Suite 340, Lincoln, NE 68583, United States
    orcid: 0000-0002-3803-0972
  - name: Klaus Ackermann
    email: Klaus.Ackermann@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0001-7693-8538
# Note that the Journal requires that a paper must begin with a
# "Summary" not an "Abstract".  This is automatically taken care
# of by the anzsauth document style.  So even though the following
# says "abstract" the heading "Summary" will appear in
# the processed version.
abstract: |
  Visual assessment of residual plots is a common approach for diagnosing linear models, but it relies on manual evaluation, which does not scale well and can lead to inconsistent decisions across analysts. The lineup protocol, which embeds the observed plot among null plots, can reduce subjectivity but requires even more human effort. In today’s data-driven world, such tasks are well-suited for automation. We present a new R package that uses a computer vision model to automate the evaluation of residual plots. An accompanying Shiny app is provided for ease of use. Given a sample of residuals, the model predicts a visual signal strength (VSS) and offers supporting information to help analysts assess model fit. 
# Note that "keywords" should not include words and phrases
# that form part of the title of the paper.
# Also keywords (with the exception of proper names and
# certain abbreviations that conventionally appear all in
# capital letters) should *not* be capitalised.
keywords: [initial data analysis, statistical graphics, data visualization, visual inference, computer vision, machine learning, hypothesis testing, regression analysis, model diagnostics]
bibliography: bibliography.bib  
#nocite: |
---

>>>>

---
title: "Automated Residual Plot Assessment with the R Package autovi and the Shiny App autovi.web"
format:
  quarto-anzjs-pdf:
    keep-tex: true  
    shorttitle: "Automated Residual Plot Assessment"
    double-space: true
    line-numbers: true
    # You can specify extra preamble in the file below.
    include-in-header:
      - header.tex
date: today
date-format: "YYYY" # used to define volumeyear
author:
  - name: Weihao Li
    email: patrick.li@anu.edu.au
    corresponding: true
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
      - name: The Australian National University
        department: Biological Data Science Institute
        address: 46 Sullivan's Creek Road, ACT 2600, Australia
    orcid: 0000-0003-4959-106X
  - name: Dianne Cook
    email: dicook@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0002-3813-7155
  - name: Emi Tanaka
    email: emi.tanaka@anu.edu.au
    affiliations: 
      - name: The Australian National University
        department: Biological Data Science Institute
        address: 46 Sullivan's Creek Road, ACT 2600, Australia
    orcid: 0000-0002-1455-259X
  - name: Susan VanderPlas
    email: emi.tanaka@anu.edu.au
    affiliations:
      - name: University of Nebraska
        department: Department of Statistics
        address: Hardin Hall, 3310 Holdrege St Suite 340, Lincoln, NE 68583, United States
    orcid: 0000-0002-3803-0972
  - name: Klaus Ackermann
    email: Klaus.Ackermann@monash.edu 
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: Wellington Road, VIC 3800, Australia
    orcid: 0000-0001-7693-8538
# Note that the Journal requires that a paper must begin with a
# "Summary" not an "Abstract".  This is automatically taken care
# of by the anzsauth document style.  So even though the following
# says "abstract" the heading "Summary" will appear in
# the processed version.
abstract: |
  Visual assessment of residual plots is a common approach for diagnosing linear models, but it relies on manual evaluation, which does not scale well and can lead to inconsistent decisions across analysts. The lineup protocol, which embeds the observed plot among null plots, can reduce subjectivity but requires even more human effort. In today’s data-driven world, such tasks are well-suited for automation. We present a new R package that uses a computer vision model to automate the evaluation of residual plots. An accompanying Shiny app is provided for ease of use. Given a sample of residuals, the model predicts a visual signal strength (VSS) and offers supporting information to help analysts assess model fit. 
# Note that "keywords" should not include words and phrases
# that form part of the title of the paper.
# Also keywords (with the exception of proper names and
# certain abbreviations that conventionally appear all in
# capital letters) should *not* be capitalised.
keywords: [initial data analysis, statistical graphics, data visualization, visual inference, computer vision, machine learning, hypothesis testing, regression analysis, model diagnostics]
bibliography: bibliography.bib  
#nocite: |
---

### 2025-05-12 autovi_paper.tex
update address and email

\author{
Weihao Li\addressnum{1},
Dianne Cook\addressnum{1},
Emi Tanaka\addressnum{2},
Susan VanderPlas\addressnum{3} and
Klaus Ackermann\addressnum{1}
}
\affiliation{
Monash University,
The Australian National University and
University of Nebraska
}

\address{
\addressnum{1} Department of Econometrics and Business
Statistics, Monash University, Wellington Road, VIC 3800, Australia\\
\addressnum{2} Biological Data Science Institute, The Australian
National University, 46 Sullivan's Creek Road, ACT 2600, Australia\\
\addressnum{3} Department of Statistics, University of Nebraska, Hardin
Hall, 3310 Holdrege St Suite 340, Lincoln, NE 68583, United States\\
\hspace*{1ex} Email: \texttt{weihao.li@monash.edu}    
}

>>>>

\author{
Weihao Li\addressnum{1, 2},
Dianne Cook\addressnum{1},
Emi Tanaka\addressnum{2},
Susan VanderPlas\addressnum{3} and
Klaus Ackermann\addressnum{1}
}
\affiliation{
Monash University,
The Australian National University and
University of Nebraska
}

\address{
\addressnum{1} Department of Econometrics and Business
Statistics, Monash University, Wellington Road, VIC 3800, Australia\\
\addressnum{2} Biological Data Science Institute, The Australian
National University, 46 Sullivan's Creek Road, ACT 2600, Australia\\
\addressnum{3} Department of Statistics, University of Nebraska, Hardin
Hall, 3310 Holdrege St Suite 340, Lincoln, NE 68583, United States\\
\hspace*{1ex} Email: \texttt{patrick.li@anu.edu.au}    
}
